---
title: "R Notebook"
output: html_notebook
---

```{r}
demographics <- read.table(file = "/Users/dc/test_stuff/analysis_factor/RapleafCompetition-master/data/demographics.tsv", sep='\t', header=TRUE)
head(demographics)
```


```{r}
library(dplyr)
library(lattice)
table(factor(demographics$home_market_value))

```
```{r}
order<-c("1k-25k","25k-50k","50k-75k","100k-150k","150k-200k","200k-250k","250k-300k","300k-350k","350k-500k","500k-1mm","1mm+")
h<-histogram(factor(demographics$home_market_value,levels=order))
h
```
```{r}
h<-histogram(factor(demographics$home_market_value))
h
```


```{r}

train <-read.csv("/Users/dc/test_stuff/analysis_factor/RapleafCompetition-master/data/training.csv",header=TRUE)
head(train)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

