---
title: "Birth Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
#install.packages("car")
#install.packages("lm.beta")
library(car)
library(lm.beta)
library(tidyverse)
library(dplyr)
```

<h1>Do not put all Xn variables in LM!! Why this is not a good
idea?</h1>

```{r}
BIRTH<-read.csv("/Users/dc/test_stuff/analysis_factor/datasets_irc_workshop/BIRTH.csv")
colnames(BIRTH)
```

APGAR10 colunn is not an integer. APGAR5, APGAR10 are infant health scores

bfacil: all 1, drop this
mage:age, 1-48
restatus: 
mrace:
married:0/1
unmarried:0/1 opposite of married column. 
meduc:
fage
frace"
feduc"
priorlive: previous births, if have 2 past births then this is 3. 
priordead: previous deaths(verify)
birth_order_cat"
last_preg"
precare"
precare_cat"
previs"
previs_rec"
cig_0"
cig_1"
cig_2"
cig_3"
cig_prior"
cig_1st"
cig_2nd"
cig_3rd"
bmi"
bmi_cat"
pre_wgt"
del_wgt"
wtgain"
wtgain_cat"
deliver_method"
attendant"
payment"
apgar5"
gestation"
gestrec10"
gest_cat2"
down_syn"
sex"
wic"
bfed"
cig_rec"
bwt"
bwt_pnds"
```{r}
sapply(BIRTH, function(df_numeric) range(df_numeric))
```
```{r}
get_numeric_cols<-select_if(BIRTH, is.numeric)
ncol(get_numeric_cols) #there should be one less column which is not numeric, APGAR10
ncol(BIRTH)
```


```{r}
sapply(BIRTH,class)
```

Birth columns:

```{r}
num=0
get_integer_cols<-function(df){
  for(i in 1:ncol(df)) {
    #print(colnames(df[i]))
    if(sapply(df[i],class)=="integer" || sapply(df[i],class)=="numeric"){
      num=num+1
      print(paste("col",i,":",colnames(df[i])))
    }
  }
  print(paste("num:",num))
  print(paste("num cols not integer or numeric:",ncol(birth)-num))
}
get_integer_cols(BIRTH)
```

```{r}
#rescaling births
BIRTH$gestation_days<-BIRTH$gestation*7
head(BIRTH$bwt,n=10)
head(BIRTH$bwt_pnds,n=10)
head(BIRTH$gestation,n=10)
head(BIRTH$gestation_days,n=10)
```

```{r}
#she ran a lm with only birth weight pounds
m_ges<-lm(BIRTH$bwt_pnds~BIRTH$gestation)
m_last_preg<-lm(BIRTH$bwt_pnds~BIRTH$last_preg)
m_bmi<-lm(BIRTH$bwt_pnds~BIRTH$bmi)
m_pre_wgt<-lm(BIRTH$bwt_pnds~BIRTH$pre_wgt)
m_del_wgt<-lm(BIRTH$bwt_pnds~BIRTH$del_wgt)
m_wtgain<-lm(BIRTH$bwt_pnds~BIRTH$wtgain)
m_bwt<-lm(BIRTH$bwt_pnds~BIRTH$bwt)
m_ges_res <- resid(m_ges)
m_last_preg_res <- resid(m_last_preg)
m_bmi_res<-resid(m_bmi)
m_pre_wgt_res<-resid(m_pre_wgt)
m_del_wgt_res<-resid(m_del_wgt)
m_wtgain_res<-resid(m_wtgain)
m_bwt_res<-resid(m_bwt)
qqnorm(m_ges_res)
qqline(m_ges_res)
qqnorm(m_last_preg_res)
qqline(m_last_preg_res)
qqnorm(m_bmi_res)
qqline(m_bmi_res)
qqnorm(m_pre_wgt_res)
qqline(m_pre_wgt_res)
qqnorm(m_del_wgt_res)
qqline(m_del_wgt_res)
qqnorm(m_wtgain_res)
qqline(m_wtgain_res)
qqnorm(m_bwt_res)
qqline(m_bwt_res)

```

OK the m_bwt_res\<-resid(m_bwt) model which is birth weight in pounds
regressed on birth weight in ounces shows a lot of skew. This is funny.
Is same as y=Constant\*y. What is this regression? This matches the
gestation is the first one, seems the residuals are normal in this
predictor.

Run a regression with 1 predictor. Correlation and cov test

```{r}
cor(BIRTH$bwt_pnds, BIRTH$gestation)
cov(BIRTH$bwt_pnds, BIRTH$gestation)
```

What is the difference between the correlation and covariance?
Correlation close to 1 or -1 means highly correlated and a 0 means not
correlated. Looks like halfway so correlated

```{r}
#these are the same variable. 
cor(BIRTH$bwt, BIRTH$bwt_pnds)
cov(BIRTH$bwt, BIRTH$bwt_pnds)
```

```{r}

corr_data<-BIRTH %>%
  select(gestation,last_preg, bmi, pre_wgt, del_wgt, wtgain, bwt, bwt_pnds)
cor(corr_data)
```

When you see a 1.00000 or high value off the diagonal. .93 is high!!!
del_wgt and pre_wgt same. pre_wgt and bmi same Those 3 closely
correlated. Gestattion not related to those 3. wt_gain also good like
gestation. bwt and bwt_pnds same. Go back and look at residuals of
wt_gain. wg_gain has bad residuals. Funny.

Covariance? Very large, not bounded. Hard to tell what is large and not.

The correlation doesnt really tell you anything. Not a 1 strongly
correlated and not a 0 not corrrelated. Same with covariance, will be
smaller or bigger if we change the units. cov dependent on magnitude of
units

```{r}
plot(BIRTH$gestation,BIRTH$bwt_pnds,type="p",ylab="Birth Weight in Pounds",xlab="Gestation in Weeks")
m_bwt_pnds_ges <-lm(BIRTH$bwt_pnds ~ BIRTH$gestation)
abline(m_bwt_pnds_ges)
```

We have some outliers which look funny. We have a workflow to identify
these. IS corr independent of units? We can verify after scaling and
centering!

```{r}
#regression diagnostics
#VIF for multicollineary only makes sense with multiple predictors
res_m_bwt_pnds_ges<-resid(m_bwt_pnds_ges)
qqnorm(res_m_bwt_pnds_ges)
qqline(res_m_bwt_pnds_ges)
skewness(res_m_bwt_pnds_ges)
kurtosis(res_m_bwt_pnds_ges)
jarque.test(res_m_bwt_pnds_ges)
shapiro.test(res_m_bwt_pnds_ges)
```

```{r}
spreadLevelPlot(m_bwt_pnds_ges)

```

```{r}
ncvTest(m_bwt_pnds_ges)
```

```{r}
outlierTest(m_bwt_pnds_ges)
```

```{r}
plotdb<-dfbetaPlots(m_bwt_pnds_ges, id.n=6)
```

We see 4,45, 1195 returned as top 3 or topN as top outliers

```{r}
influencePlot(m_bwt_pnds_ges)
```

```{r}
#this shouldnt be different from above! 
plot(BIRTH$gestation_days,BIRTH$bwt,type="p",ylab="Birth Weight in Ounces",xlab="Gestation in Days")
abline(lm(BIRTH$bwt ~ BIRTH$gestation_days))
```

```{r}
model5<-lm(bwt_pnds~gestation,data=BIRTH)
summary(model5)
```

```{r}
model6<-lm(bwt~gestation_days,data=BIRTH)
summary(model6)
```

```{r}
NLSY<-nlsy
model7<-lm(PCS2000~CESD2000Total+Education2000+NumberBioStepAdoptChildHH2000+MCS2000,data=NLSY)
model7.1<-lm.beta(model7)
summary(model7.1)

```

```{r}
#missing sqrt term had to fix
BIRTH$ln_last_preg<-log(BIRTH$last_preg)
BIRTH$inv_last_preg<-1/(BIRTH$last_preg)
BIRTH$sqrt_last_preg<-sqrt(BIRTH$last_preg)
BIRTH$ln_pre_wgt<-log(BIRTH$pre_wgt)
head(BIRTH$last_preg,n=10)
head(BIRTH$ln_last_pre,n=10)
head(BIRTH$inv_last_preg,n=10)
head(BIRTH$sqrt_last_preg,n=10)
head(BIRTH$ln_pre_wgt,n=10)

```

```{r}
plot(BIRTH$mage,BIRTH$last_preg,type="p",ylab="last_preg",xlab="Mother's Age")
abline(lm(BIRTH$last_preg ~ BIRTH$mage))

plot(BIRTH$mage,BIRTH$sqrt_last_preg,type="p",ylab="sqrt_last_preg",xlab="Mother's Age")
abline(lm(BIRTH$sqrt_last_preg ~ BIRTH$mage))

plot(BIRTH$mage,BIRTH$ln_last_preg,type="p",ylab="ln_last_preg",xlab="Mother's Age")
abline(lm(BIRTH$ln_last_preg ~ BIRTH$mage))

##histograms of untransformed, square root, and natural log of months since last pregnancy
hist(BIRTH$last_preg)
hist(BIRTH$sqrt_last_preg)
hist(BIRTH$ln_last_preg)

```

```{r}
##Square Root transformation on Y
model8<-lm(sqrt_last_preg~mage+pre_wgt,data=BIRTH)
summary(model8)

##Natural log transformation on Y
model9<-lm(ln_last_preg~mage+pre_wgt,data=BIRTH)
summary(model9)

##Natural log transformation on X
model10<-lm(last_preg~mage+ln_pre_wgt,data=BIRTH)
summary(model10)

plot(BIRTH$ln_pre_wgt,BIRTH$last_preg,type="p",ylab="Months Since Last Pregnancy",xlab="ln_pre_wgt")
abline(lm(BIRTH$last_preg ~ BIRTH$ln_pre_wgt))

##Natural log transformation on both Y and X
model11<-lm(ln_last_preg~mage+ln_pre_wgt,data=BIRTH)
summary(model11)

plot(BIRTH$ln_pre_wgt,BIRTH$ln_last_preg,type="p",ylab="ln_last_preg",xlab="ln_pre_wgt")
abline(lm(BIRTH$ln_last_preg ~ BIRTH$ln_pre_wgt))


```
