---
title: "R Notebook"
output: html_notebook
---


```{r}
install.packages("tidyverse")
install.packages("ggpubr")
install.packages("rstatix")

library(tidyverse)
library(ggpubr)
library(rstatix)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
```{r}
install.packages("datarium")
library(datarium)
data("selfesteem", package = "datarium")
head(selfesteem, 3)
```
```{r}
selfesteem <- selfesteem %>%
  gather(key = "time", value = "score", t1, t2, t3) %>%
  convert_as_factor(id, time)
head(selfesteem, 3)
```

```{r}
selfesteem %>%
  group_by(time) %>%
  get_summary_stats(score, type = "mean_sd")
```
```{r}
bxp <- ggboxplot(selfesteem, x = "time", y = "score", add = "point")
bxp
```
```{r}
selfesteem %>%
  group_by(time) %>%
  identify_outliers(score)
```

